---
title: "Plot1Face Tutorial"
author: "Julie D. White"
date: "May 14, 2018"
output: html_document
always_allow_html: yes
---

This tutorial is set to accompany the paper by Zaidi & White, et al. "Facial masculinity does not appear to be a condition-dependent male ornament in humans and does not reflect MHC heterozygosity" to provide the user with a way of visualizing dense heatmaps on facial scans.  

The Plot1Face.R function in written by Tomas Gonzalez-Zarzar. Follow him at: https://github.com/tomszar  

Load necessary libraries
``` {r}
require(plotly)
require(mixOmics)
require(data.table)
```

Read in provided obj file. The contains information for the template face on top of which heatmaps can be visualized. The file contains 7,150 landmark vertices and accompanying triangle facet information.
```{r}
RefScan <- read.table("RefScan.obj", sep = "\t", header = F)
head(RefScan)
```

Split the information into vertices and facets matrices. 
```{r}
RefScan_Vertices <- as.matrix(RefScan[which(RefScan$V1 == "v"), 2:4])
RefScan_Facets <- as.matrix(RefScan[which(RefScan$V1 == "f"), 2:4])
head(RefScan_Vertices)
head(RefScan_Facets)
```

Initialize Plot1Face function from source file
```{r}
source('../Scripts/Plot1Face.R')
```

Example of plotting a face in greyscale, with no heatmap information.
```{r}
Plot1Face(vertices = RefScan_Vertices, facets = RefScan_Facets, colormap = NULL, title = "Example greyscale face")
```

Examples using results from Zaidi and White, et al. paper.

Read in FM_QL file
```{r}
qlmasc<-fread("../Dataset/qlmasc_1233_noprop_03292018.txt",header=F)
```


Plot facial masculinity of a face which has greater overall facial masculinity than the female consensus face
```{r}
#read dataframe containing IDs and overall facial masculinity measurements
eurofam<-read.table("../Dataset/euro_1233_masc_het_03292018.dat",sep="\t",header=T)
#determine the index in the ql file which corresponds to the ID you are looking for
id1.index<-sample(which(eurofam$avg.masc.unit>2),1)
#isolate the FM_ql values for this person
id1.ql<-t(qlmasc[id.index,])
Plot1Face(vertices = RefScan_Vertices, facets = RefScan_Facets, colormap = id1.ql, title = "FM_ql")
```
Plot geometric sexual dimorphism for each vertex and Cohen's D estimate for each vertex 
(or any other statistic e.g. effect size of predictor/p-value etc.)
```{r}
#Geometric sexual dimorphism
ql_gsd <- read.table("ql_gsd_03292018.txt", sep = "\t", header = T)
head(ql_gsd)
```

```{r}
#Cohen's D
ql_sex_cohenD <- read.table("../Summary_dat/ql_sex_cohenD_03292018.txt", sep = "\t", header = T)
head(ql_sex_cohenD)
```

Add colors to the face representing the sexual dimorphism at each vertex.
```{r}
Plot1Face(vertices = RefScan_Vertices, facets = RefScan_Facets, colormap = ql_gsd$x, title = "Geometric sexual dimorphism")
```

Add colors to the face representing the Cohen's D estimate for sex difference at each vertex.
```{r}
Plot1Face(vertices = RefScan_Vertices, facets = RefScan_Facets, colormap = ql_sex_cohenD$x, title = "Cohen's D estimate for sexual dimorphism")
```

The color scale can be adjusted to a specified minimum and maximum
Otherwise it defaults to the minimum and maximum of the values provided
```{r}
Plot1Face(vertices = RefScan_Vertices, facets = RefScan_Facets, colormap = ql_sex_cohenD$x, title = "Cohen's D estimate for sexual dimorphism",color.min = 0,color.max=3)
```

